### 2017-01-24 homework/week-1/main.tex
Adding references and some changes to homework 1

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath} 
\usepackage{amsthm}
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\usepackage{float}
\setlength\parindent{0pt}

@@

@@

Let \(\chi\) refer to the coloring generated by using the greedy algorithm with degree sequencing heuristic. Let \(\chi^{*}\) be the optimal coloring.

@@

\begin{figure}[H]
\centering
\includegraphics[scale=0.5]{graph-1.png}
\caption{Uncolored original graph \(G\)}
\end{figure}

By applying the greedy algorithm plus degree sequencing heuristic, we get the following coloring. The numbers indicate the ordering of vertices before applying the greedy algorithm. Any vertex of the same degree got assigned arbitrarily. This results in \(\chi(G) = 3\).

\begin{figure}[H]
\centering
\includegraphics[scale=0.5]{graph-2.png}
\caption{Coloring from applying the greedy plus degree sequencing heuristic}
\end{figure}

We can see that this is a bipartite graph and thus is \emph{2-colorable}. This means \(\chi^{*}(G) = 2\). Thus, this is an example graph that satisfies conditions \((1)\) and \((2)\). The optimal coloring is shown below.

\begin{figure}[H]
\centering
\includegraphics[scale=0.5]{graph-3.png}
\caption{Optimal coloring of graph \(G\)}
\end{figure}

\newpage

@@

To do so, let's create another example that satisfies conditions \((1)\) and \((2)\) from above. After doing some research into bipartite graphs, I learned that \emph{crown graphs} are excellent at showing how bad greedy heuristics can be.

\begin{definition}[Crown Graph]
A crown graph is a complete bipartite graph where the edges of a perfect matching have been removed.
\end{definition}

@@

\begin{figure}[H]
\centering
\includegraphics[scale=0.40]{graph-4.png}
\caption{Uncolored original graph \(H\)}
\end{figure}

@@

\begin{figure}[H]
\centering
\includegraphics[scale=0.40]{graph-5.png}
\caption{Worst-case coloring of \(H\) using either greedy heuristic}
\end{figure}

\newpage

In Figure 5, using either greedy heuristic with this ordering, we get \(\chi(H) = 5\). This gives us \(\frac{|V|}{2}\) colors. This is the worst-case for a crown graph, but this graph also demonstrates the upper bounds for these greedy algorithms. \newline

@@

We can see that \(\chi(H) = \Delta(H) + 1\). This is very easy to see in a \emph{complete} graph, where all vertices are connected to every other vertex. This means all vertices have degree \(|V| - 1\). Thus, every time we color a node, a new color is needed. And since we have \(\Delta(P) = |V| - 1\) and \(|V|\) vertices, we will need \(\Delta(P) + 1\) colors. This is proved in \emph{Brooks' Theorem}.

@@

Overall, the greedy and the greedy plus degree sequencing heuristics can produce some very bad results. In graph \(H\), at the worst case, the greedy heuristic produces \(\chi(H) = 5\) when \(\chi^{*}(H) = 2\) as it is bipartite. This is shown below.

\begin{figure}[H]
\centering
\includegraphics[scale=0.40]{graph-6.png}
\caption{Optimal coloring of \(H\)}
\end{figure}

>>>>

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage[left=1in,right=1in,top=0.9in,bottom=0.9in]{geometry}
\usepackage{graphicx}
\usepackage{float}
\setlength\parindent{0pt}

@@

\section*{Heuristics}
In this report, I will use two vertex coloring heuristics as defined in class:
\begin{itemize}
\item \textbf{Heuristic A}: Greedy algorithm
\item \textbf{Heuristic B}: Greedy algorithm plus degree sequencing
\end{itemize}

Heuristic B is also known as the Welsh-Powell algorithm, which is defined in \cite{welsh}.

@@

Let \(\chi\) refer to the coloring generated by \emph{heuristic B}. Let \(\chi^{*}\) be the optimal coloring.

@@

\begin{figure}[H]
\centering
\includegraphics[scale=0.6]{graph-1.png}
\caption{Uncolored original graph \(G\)}
\end{figure}

By applying \emph{heuristic B}, we get the following coloring. The numbers indicate the ordering of vertices before applying the heuristic. Any vertex of the same degree got assigned arbitrarily. This results in \(\chi(G) = 3\).

\begin{figure}[H]
\centering
\includegraphics[scale=0.6]{graph-2.png}
\caption{Coloring from applying heuristic B}
\end{figure}

\begin{definition}[Bipartite graph]
A bipartite graph is one whose vertex set can be partitioned into two subsets X and Y such that each edge has one end in X and one end in Y
\end{definition}

We can see that this is a bipartite graph, defined above by \cite{bondymurty}. Thus, \(G\) is \emph{2-colorable}. This means \(\chi^{*}(G) = 2\). \(G\) is an example graph that satisfies conditions \((1)\) and \((2)\). The optimal coloring is shown below.

\begin{figure}[H]
\centering
\includegraphics[scale=0.6]{graph-3.png}
\caption{Optimal coloring of graph \(G\)}
\end{figure}

@@

To do so, let's create another example that satisfies conditions \((1)\) and \((2)\) from above. After doing some research into bipartite graphs, I learned that \emph{crown graphs} are excellent at showing how bad greedy heuristics can be, as shown and defined in \cite{kordecki}.

\begin{definition}[Crown Graph]
A crown graph \(C_n = (V, E)\) is an undirected graph with two sets of vertices where \(V = V_1 \cup V_2\) with an edge from \(v_{1,i}\) to \(v_{2,j}\) whenever \(i \neq j\). A crown graph can also be viewed as a complete bipartite graph where the edges of a perfect matching have been removed.
\end{definition}

@@

\begin{figure}[H]
\centering
\includegraphics[scale=0.38]{graph-4.png}
\caption{Uncolored original graph \(H\)}
\end{figure}

@@

\begin{figure}[H]
\centering
\includegraphics[scale=0.38]{graph-5.png}
\caption{Worst-case coloring of \(H\) using either greedy heuristic}
\end{figure}

In Figure 5, using either heuristic with this ordering, we get \(\chi(H) = 5\). This gives us \(\frac{|V|}{2}\) colors. This is the worst-case for a crown graph as shown in \cite{johnson}, but this graph also demonstrates the upper bounds for these heuristics. \newline

@@

We can see that \(\chi(H) = \Delta(H) + 1\). This is very easy to see in a \emph{complete} graph, where all vertices are connected to every other vertex. This means all vertices have degree \(|V| - 1\). Thus, every time we color a node, a new color is needed. And since we have \(\Delta(P) = |V| - 1\) and \(|V|\) vertices, we will need \(\Delta(P) + 1\) colors. This is proved in \emph{Brooks' Theorem}, defined below from \cite{bondymurty}.

@@

Overall, heuristic A and heuristic B can produce some very undesirable results. In graph \(H\), at the worst case, these heuristics produce \(\chi(H) = 5\) when \(\chi^{*}(H) = 2\) as it is bipartite. This is shown below.

\begin{figure}[H]
\centering
\includegraphics[scale=0.38]{graph-6.png}
\caption{Optimal coloring of \(H\)}
\end{figure}

\begin{thebibliography}{9}
\bibitem{bondymurty}
Bondy, J.A. and U.S.R. Murty [1976],
\emph{Graph Theory with Applications},
American Elsevier Publishing, New York, NY.

\bibitem{johnson}
Johnson, D. S. [1974], \emph{Worst-case behavior of graph coloring algorithms}, Proc. 5th Southeastern Conf. on Combinatorics, Graph Theory, and Computing, Utilitas Mathematicae, Winnipeg, pp. 513–527.

\bibitem{kordecki}
Kordecki, W. and A. Łyczkowska-Hanćkowiak [2016], \emph{Greedy online colouring with buffering}, arXiv preprint arXiv:1601.00252.

\bibitem{welsh}
Welsh, D. J. and Powell, M. B. [1967], \emph{An upper bound for the chromatic number of a graph and its application to timetabling problems}, The Computer Journal, 10(1), 85-86.

\end{thebibliography}
